### 2024-10-22 paper/paper.qmd
rough abstract and intro

---
title: "Race to the Ballot Box: Will Demographics Decide a Winner This Year?"
subtitle: "A Regressional Approach to Race and Political Preferences in the 2024 US Elections and Insights on the Next"
author: 
  - Sameeck Bhatia
  - Sean Chua
  - Tanmay Shinde
thanks: "Code and data are available at: [https://github.com/SameeckBhatia/2024-US-Elections](https://github.com/SameeckBhatia/2024-US-Elections)."
date: today
date-format: long
abstract: "This paper analyzes polling data for the 2024 U.S. Presidential election to predict the winner, while also capturing demographic sentiment toward the two main contenders - Donald Trump and Kamala Harris. We found that ....[NEED TO COMPLETE]. This study highlights that factors such as .... the reputation of the pollster, quality of the poll, bias, sample size, demographic area...[COME BACK AND CHANGE] are essential for analyzing polling data and thereby forecasting election outcomes. Highlighting these trends is crucial for understanding the evolving political landscape in the U.S., as they reveal how voter demographics and sentiments shape election outcomes."
format: pdf
toc: true
number-sections: true
bibliography: references.bib
---

```{r}
#| include: false
#| warning: false
#| message: false

library(tidyverse)
```

# Introduction

\textbf{Broader Context + What the Paper is About} As the 2024 U.S. Presidential election draws near, the race between former President Donald Trump and Vice President Kamala Harris has garnered widespread attention worldwide. With both candidates representing contrasting ideologies and policy priorities, appealing to distinct voter bases, understanding the public sentiment surrounding them becomes crucial in predicting the election's outcome. Polls and demographic analyses provide a lens through which we can examine the factors influencing voter preferences and gauge the political landscape as it evolves. This paper aims to analyze the 2024 election polling data to not only forecast the winner but also to capture the demographic-wise public sentiment surrounding the two contenders.

\textbf{Estimand:} The primary estimand for this analysis is the predicted support percentage for each of the two candidates.

\textbf{What was done:} We used polling data from FiveThirtyEight, which includes surveys from organizations such as YouGov, Siena/NYT, and Washington Post. We used a Bayesian approach to model the polling support percentages for both candidates, and understand how polling factors such as numeric grade, poll score, state, and sample size affect the predicted support percentage.

The results show that ... [TO-DO]. These results are important because ... [TO-DO].

The rest of the paper is structured as follows: @sec-data details the data and measurement process.@sec-model presents the model and justifies the choices made in the building of the chosen model. @sec-results presents the results, highlighting the relationship between different variables and the polling support percentage, and @sec-discussion discusses the implications of the findings for this as well as future election forecasting.

# Data {#sec-data}

## Overview

We use the statistical programming language R [@citeR].... Our data [@shelter].... Following @tellingstories, we consider...

Lorem ipsum dolor sit amet, consectetur adipiscing elit. Aliquam auctor quam et justo efficitur, sit amet eleifend elit euismod. Mauris non libero vel ligula tincidunt consequat. Aenean gravida, risus id auctor aliquam, orci leo auctor tellus, non tincidunt est arcu at lectus. Vivamus feugiat et quam ut consequat. Aenean mollis ullamcorper facilisis. Vivamus scelerisque lectus et elementum vulputate. Nulla et arcu vehicula, iaculis felis sit amet, semper justo. Suspendisse quam augue, hendrerit a tortor non, tristique feugiat sem. Curabitur vitae tortor nec ligula scelerisque rutrum. Nullam feugiat odio metus. Cras id convallis ante, ut ornare velit. Mauris turpis purus, porttitor eu leo quis, suscipit euismod ligula.

## Measurement

Lorem ipsum dolor sit amet, consectetur adipiscing elit. Aliquam auctor quam et justo efficitur, sit amet eleifend elit euismod. Mauris non libero vel ligula tincidunt consequat. Aenean gravida, risus id auctor aliquam, orci leo auctor tellus, non tincidunt est arcu at lectus. Vivamus feugiat et quam ut consequat. Aenean mollis ullamcorper facilisis. Vivamus scelerisque lectus et elementum vulputate. Nulla et arcu vehicula, iaculis felis sit amet, semper justo. Suspendisse quam augue, hendrerit a tortor non, tristique feugiat sem. Curabitur vitae tortor nec ligula scelerisque rutrum. Nullam feugiat odio metus. Cras id convallis ante, ut ornare velit. Mauris turpis purus, porttitor eu leo quis, suscipit euismod ligula.

## Outcome variables

Lorem ipsum dolor sit amet, consectetur adipiscing elit. Aliquam auctor quam et justo efficitur, sit amet eleifend elit euismod. Mauris non libero vel ligula tincidunt consequat. Aenean gravida, risus id auctor aliquam, orci leo auctor tellus, non tincidunt est arcu at lectus. Vivamus feugiat et quam ut consequat. Aenean mollis ullamcorper facilisis. Vivamus scelerisque lectus et elementum vulputate. Nulla et arcu vehicula, iaculis felis sit amet, semper justo. Suspendisse quam augue, hendrerit a tortor non, tristique feugiat sem. Curabitur vitae tortor nec ligula scelerisque rutrum. Nullam feugiat odio metus. Cras id convallis ante, ut ornare velit. Mauris turpis purus, porttitor eu leo quis, suscipit euismod ligula.

## Predictor variables

Lorem ipsum dolor sit amet, consectetur adipiscing elit. Aliquam auctor quam et justo efficitur, sit amet eleifend elit euismod. Mauris non libero vel ligula tincidunt consequat. Aenean gravida, risus id auctor aliquam, orci leo auctor tellus, non tincidunt est arcu at lectus. Vivamus feugiat et quam ut consequat. Aenean mollis ullamcorper facilisis. Vivamus scelerisque lectus et elementum vulputate. Nulla et arcu vehicula, iaculis felis sit amet, semper justo. Suspendisse quam augue, hendrerit a tortor non, tristique feugiat sem. Curabitur vitae tortor nec ligula scelerisque rutrum. Nullam feugiat odio metus. Cras id convallis ante, ut ornare velit. Mauris turpis purus, porttitor eu leo quis, suscipit euismod ligula.

# Model {#sec-model}

Lorem ipsum dolor sit amet, consectetur adipiscing elit. Aliquam auctor quam et justo efficitur, sit amet eleifend elit euismod. Mauris non libero vel ligula tincidunt consequat. Aenean gravida, risus id auctor aliquam, orci leo auctor tellus, non tincidunt est arcu at lectus. Vivamus feugiat et quam ut consequat. Aenean mollis ullamcorper facilisis. Vivamus scelerisque lectus et elementum vulputate. Nulla et arcu vehicula, iaculis felis sit amet, semper justo. Suspendisse quam augue, hendrerit a tortor non, tristique feugiat sem. Curabitur vitae tortor nec ligula scelerisque rutrum.

## Model set-up

Lorem ipsum dolor sit amet, consectetur adipiscing elit. Aliquam auctor quam et justo efficitur, sit amet eleifend elit euismod. Mauris non libero vel ligula tincidunt consequat. Aenean gravida, risus id auctor aliquam, orci leo auctor tellus, non tincidunt est arcu at lectus. Vivamus feugiat et quam ut consequat. Aenean mollis ullamcorper facilisis. Vivamus scelerisque lectus et elementum vulputate. Nulla et arcu vehicula, iaculis felis sit amet, semper justo. Suspendisse quam augue, hendrerit a tortor non, tristique feugiat sem. Curabitur vitae tortor nec ligula scelerisque rutrum. Nullam feugiat odio metus. Cras id convallis ante, ut ornare velit. Mauris turpis purus, porttitor eu leo quis, suscipit euismod ligula.

# Results {#sec-results}

Lorem ipsum dolor sit amet, consectetur adipiscing elit. Aliquam auctor quam et justo efficitur, sit amet eleifend elit euismod. Mauris non libero vel ligula tincidunt consequat. Aenean gravida, risus id auctor aliquam, orci leo auctor tellus, non tincidunt est arcu at lectus. Vivamus feugiat et quam ut consequat. Aenean mollis ullamcorper facilisis. Vivamus scelerisque lectus et elementum vulputate. Nulla et arcu vehicula, iaculis felis sit amet, semper justo. Suspendisse quam augue, hendrerit a tortor non, tristique feugiat sem. Curabitur vitae tortor nec ligula scelerisque rutrum. Nullam feugiat odio metus. Cras id convallis ante, ut ornare velit. Mauris turpis purus, porttitor eu leo quis, suscipit euismod ligula.

Lorem ipsum dolor sit amet, consectetur adipiscing elit. Aliquam auctor quam et justo efficitur, sit amet eleifend elit euismod. Mauris non libero vel ligula tincidunt consequat. Aenean gravida, risus id auctor aliquam, orci leo auctor tellus, non tincidunt est arcu at lectus. Vivamus feugiat et quam ut consequat. Aenean mollis ullamcorper facilisis. Vivamus scelerisque lectus et elementum vulputate. Nulla et arcu vehicula, iaculis felis sit amet, semper justo. Suspendisse quam augue, hendrerit a tortor non, tristique feugiat sem. Curabitur vitae tortor nec ligula scelerisque rutrum. Nullam feugiat odio metus. Cras id convallis ante, ut ornare velit. Mauris turpis purus, porttitor eu leo quis, suscipit euismod ligula.

# Discussion {#sec-discussion}

## First discussion point {#sec-first-point}

Lorem ipsum dolor sit amet, consectetur adipiscing elit. Aliquam auctor quam et justo efficitur, sit amet eleifend elit euismod. Mauris non libero vel ligula tincidunt consequat. Aenean gravida, risus id auctor aliquam, orci leo auctor tellus, non tincidunt est arcu at lectus. Vivamus feugiat et quam ut consequat. Aenean mollis ullamcorper facilisis. Vivamus scelerisque lectus et elementum vulputate. Nulla et arcu vehicula, iaculis felis sit amet, semper justo. Suspendisse quam augue, hendrerit a tortor non, tristique feugiat sem. Curabitur vitae tortor nec ligula scelerisque rutrum. Nullam feugiat odio metus. Cras id convallis ante, ut ornare velit. Mauris turpis purus, porttitor eu leo quis, suscipit euismod ligula.

## Second discussion point

Lorem ipsum dolor sit amet, consectetur adipiscing elit. Aliquam auctor quam et justo efficitur, sit amet eleifend elit euismod. Mauris non libero vel ligula tincidunt consequat. Aenean gravida, risus id auctor aliquam, orci leo auctor tellus, non tincidunt est arcu at lectus. Vivamus feugiat et quam ut consequat. Aenean mollis ullamcorper facilisis. Vivamus scelerisque lectus et elementum vulputate. Nulla et arcu vehicula, iaculis felis sit amet, semper justo. Suspendisse quam augue, hendrerit a tortor non, tristique feugiat sem. Curabitur vitae tortor nec ligula scelerisque rutrum. Nullam feugiat odio metus. Cras id convallis ante, ut ornare velit. Mauris turpis purus, porttitor eu leo quis, suscipit euismod ligula.

## Third discussion point

Lorem ipsum dolor sit amet, consectetur adipiscing elit. Aliquam auctor quam et justo efficitur, sit amet eleifend elit euismod. Mauris non libero vel ligula tincidunt consequat. Aenean gravida, risus id auctor aliquam, orci leo auctor tellus, non tincidunt est arcu at lectus. Vivamus feugiat et quam ut consequat. Aenean mollis ullamcorper facilisis. Vivamus scelerisque lectus et elementum vulputate. Nulla et arcu vehicula, iaculis felis sit amet, semper justo. Suspendisse quam augue, hendrerit a tortor non, tristique feugiat sem. Curabitur vitae tortor nec ligula scelerisque rutrum. Nullam feugiat odio metus. Cras id convallis ante, ut ornare velit. Mauris turpis purus, porttitor eu leo quis, suscipit euismod ligula.

## Weaknesses and next steps

Lorem ipsum dolor sit amet, consectetur adipiscing elit. Aliquam auctor quam et justo efficitur, sit amet eleifend elit euismod. Mauris non libero vel ligula tincidunt consequat. Aenean gravida, risus id auctor aliquam, orci leo auctor tellus, non tincidunt est arcu at lectus. Vivamus feugiat et quam ut consequat. Aenean mollis ullamcorper facilisis. Vivamus scelerisque lectus et elementum vulputate. Nulla et arcu vehicula, iaculis felis sit amet, semper justo. Suspendisse quam augue, hendrerit a tortor non, tristique feugiat sem. Curabitur vitae tortor nec ligula scelerisque rutrum. Nullam feugiat odio metus. Cras id convallis ante, ut ornare velit. Mauris turpis purus, porttitor eu leo quis, suscipit euismod ligula.

\newpage

\appendix

# Appendix {.unnumbered}

\newpage

# References